---
title: "Rossman Store Sales "
subtitle: "Group 1 - Exploratory Data Analysis"
author: '1. Kang Jun Han Brandon, 2.Kenny Sim Jun Hong, 3. Lim Yong Chuan,  4. Tan Soon Wei,  5. Teo Xiangquan Martin'
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: 
    tufte_variant: "default"
    self_contained: yes
---
```{r libraries, include = F}
# Install & load relevant libraries. 
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, 
               ggplot2,
               lubridate, 
               plotly,
               kableExtra,
               Knitr)

# Create html_df for later stlying
html_df <- function(x){
  kable(x) %>%
   kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
}
```

# Importing and cleaning data
## 1. Import files
``` {r Import, eval = T}
# Import train.csv, test.csv and store.csv
train <- read.csv("train.csv", stringsAsFactors = F)
test <- read.csv("test.csv", stringsAsFactors = F)
store <- read.csv("store.csv", stringsAsFactors = F)
```

## 2. Data Structure
```{r Dimensions Check, include = F}
 str(train)
 str(test)
 str(store)
 dim(train)
 dim(test)
 dim(store)
```
 
```{r Data Structure, eval = T, echo= F }
# Rows and columns of each file
matrix(c("1017209", "9", "41088", "8", "1115", "10"),ncol=2, byrow= TRUE)%>% as.data.frame() %>% `row.names<-`(c("Train", "Test", "Store")) %>% `colnames<-`(c("No. of Rows", "No. of Columns")) %>% html_df
```

## 3. Check for NA 
### NA values for train ###   
```{r NA values - train, eval = T, echo = F}
# Train 
train %>% is.na() %>% colSums() %>% data.frame() %>% `colnames<-`("No. of NAs") %>% html_df
```


### NA values for test ###   
``` {r NA values - test, eval = T, echo = F}
# Test 
test %>% is.na() %>% colSums() %>% data.frame() %>% `colnames<-`("No. of NAs") %>% html_df %>% row_spec(5, bold = T, color = "red")
```


_Observations_:

+ Variable 'Open' should have only two possible values (Open = 1 or Closed = 0), so the 11 NA's should be changed to either 1 or 0.
+ If Open is = 1, but we assume = 0, the error score will increase because of misprediction.
+ If Open is = 0, but we assume = 1, then there's no penalty in scoring as closed stores with 0 sales are not considered in scoring. 

Hence, we will impute 1 into the NA values for the 'Open' variable in the test dataset. 

### NA values for store ###   
```{r NA values - Store, eval = T, echo = F}
# Store
store %>% is.na() %>% colSums() %>% data.frame() %>% `colnames<-`("No. of NAs") %>% html_df
```



## 4. Imputing missing values for test
``` {r Impute, eval = T}
# a. Retrieve records with Open = NA
test %>% filter(is.na(Open)) %>% html_df()

# b. Replace NA with Open = 1
test <- test %>% mutate(Open = replace(Open, is.na(Open),1))

# c. Check if NA has been replaced:
sum(is.na(test$Open))
```



## 5. Convert data types

+ Insert explanation for the type conversion here

``` {r Converting data, eval = T}
#a. Train
train <- train %>% mutate(
  DayOfWeek                 = as.factor(DayOfWeek),
  Date                      = as.Date(Date),
  Open                      = as.factor(Open),
  Promo                     = as.factor(Promo), 
  StateHoliday              = as.factor(StateHoliday),   # Has 4 values!
  SchoolHoliday             = as.factor(SchoolHoliday),
  Day                       = as.integer(format(train$Date, "%d")), # New variable 1
  Month                     = as.integer(format(train$Date, "%m")), # New variable 2
  Year                      = as.integer(format(train$Date, "%Y"))) # New variable 3
str(train)

#b. Test
test <- test %>% mutate(
  DayOfWeek                 = as.factor(DayOfWeek),
  Date                      = as.Date(Date),
  Open                      = as.factor(Open),
  Promo                     = as.factor(Promo),
  StateHoliday              = as.factor(StateHoliday),   # Only 2 values! What're the state holidays?
  SchoolHoliday             = as.factor(SchoolHoliday),
  Day                       = as.integer(format(test$Date, "%d")),  # New variable 1
  Month                     = as.integer(format(test$Date, "%m")),  # New variable 2
  Year                      = as.integer(format(test$Date, "%Y")))  # New variable 3
str(test)

#c. Store
store <- store %>% mutate(
  StoreType                 = as.factor(StoreType),
  Assortment                = as.factor(Assortment),
  Promo2                    = as.factor(Promo2))
str(store)

```


# Exploratory Data Analysis
## 1. Dates
Sales data start from `r min(train$Date)` to `r max(train$Date)`, which spans a total of `r max(train$Date) - min(train$Date)` days or 2 years 7 months.

## 2. Sales - By day of week (Dependent Variable)

__Observations__:
Why are Sunday's sales so low? To check further
```{r Sales - DayofWeek, eval = T }
ggplot(data = train, aes (x= DayOfWeek, y= Sales)) +
geom_bar(stat = "identity")

count <- train %>%   count(DayOfWeek, sort = T)
count
unique(train$DayOfWeek)

```
## 3

## Reserved . Skewedness of data
```{r, eval = T}
#group revenues by quarter
train <- train %>%
          mutate(quarters = paste0(substr(Date,1,4),quarters(as.Date(Date))))

train <- train %>% 
          group_by(quarters) %>%
          mutate(revtq = sum(Sales,na.rm = TRUE)) %>%
          ungroup()
```


          
