---
title: "Rossman Store Sales "
subtitle: "Group 1 - Exploratory Data Analysis"
author: '1. Kang Jun Han Brandon, 2.Kenny Sim Jun Hong, 3. Lim Yong Chuan,  4. Tan Soon Wei,  5. Teo Xiangquan Martin'
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: 
    tufte_variant: "default"
    self_contained: yes
---
```{r libraries, include = FALSE}
# Install & load relevant libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, readr, lubridate, pryr)
```

# Data Cleaning
## 1. Importing files 
Import files from train, store, and test.
``` {r read, eval = TRUE}
# a. Import train.csv
train <- read.csv("train.csv", stringsAsFactors = F)

# b. Import test.csv
test <- read.csv("test.csv", stringsAsFactors = F)

# c.  Import store.csv
store <- read.csv("store.csv", stringsAsFactors = F)

```


## 2. Check for NA values
```{r NA values, eval = TRUE}
# a. Train 
colSums(is.na(train)) 

# b. Test 
colSums(is.na(test))

# c. Store
colSums(is.na(store)) 
```

## 3. Imputing missing values for test
``` {r Impute, eval = TRUE}
# a. Retrieve records with Open = NA
test %>% filter(is.na(Open))

# b. Replace NA with Open = 1
test <- test %>% mutate(Open = replace(Open, is.na(Open),1))

# c. Check if NA has been replaced:
sum(is.na(test$Open))
```
__Explanation__:
+ If we assume


## 3. Convert data types
``` {r Converting data, eval = TRUE}
#a. Train
train <- train %>% mutate(
  DayOfWeek                 = as.factor(DayOfWeek),
  Date                      = as.Date(Date),
  Open                      = as.factor(Open),
  Promo                     = as.factor(Promo), 
  StateHoliday              = as.factor(StateHoliday),   # Has 4 values!
  SchoolHoliday             = as.factor(SchoolHoliday))
str(train)

#b. Test
test <- test %>% mutate(
  DayOfWeek                 = as.factor(DayOfWeek),
  Date                      = as.Date(Date),
  Open                      = as.factor(Open),
  Promo                     = as.factor(Promo),
  StateHoliday              = as.factor(StateHoliday),   # Only 2 values! What're the state holidays?
  SchoolHoliday             = as.factor(SchoolHoliday))
str(test)

#c. Store
store <- store %>% mutate(
  StoreType                 = as.factor(StoreType),
  Assortment                = as.factor(Assortment),
  Promo2                    = as.factor(Promo2))
str(store)

```

## 4. Skewedness of data
library(tidyverse)
library(plotly)
library(lubridate)

#group revenues by quarter
train <- train %>%
          mutate(quarters = paste0(substr(Date,1,4),quarters(as.Date(Date))))

train <- train %>% 
          group_by(quarters) %>%
          mutate(revtq = rowSums(Sales,na.rm = TRUE) %>%
          ungroup()
